---
title: "Homework 3"
author: Jingyi Zhang
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
          
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user/ order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.


##### How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

##### Make a plot

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

##### Make a table showing the three most popular items in each of the aisles.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

##### Make a table showing apples vs. ice cream

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()
```

<br />

## Problem 2

##### Load and tidy the dataset.

```{r tidy_dataset}
accel_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    day = factor(day),
    weekday = case_when(
      day == "Monday" ~ "Monday",
      day == "Tuesday" ~ "Tuesday",
      day == "Wednesday" ~ "Wednesday",
      day == "Thursday" ~ "Thursday",
      day == "Friday" ~ "Friday"),
    weekend = case_when(
      day == "Saturday" ~ "Saturday",
      day == "Sunday" ~ "Sunday")
    ) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_counts"
  )
```

This dataset contains information on five weeks of accelerometer data collected from a 63 year-old male with BMI 25 and diagnosed with congestive heart failure. After cleaning and tidying, the final dataset contains week, day id, day, weekday, weekend, minute, and activity counts columns. It has a total of `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. 

<br />

##### Create a total activity variable for each day, and create a table showing these totals.


```{r}
accel_df %>%
  group_by(day_id) %>%
  summarize(total_activity = sum(activity_counts)) %>%
  knitr::kable()
```

Describe trends?

##### Make a single-panel plot 

```{r}
accel_df %>%
  ggplot(aes(x = minute, y = activity_counts, color = day)) +
  geom_point(alpha = .7) +
  labs(
    title = "24-hour Activity Plot",
    x = "24-hour (min)",
    y = "Activity counts",
    caption = "Data from Advanced Cardiac Care Center of Columbia University Medical Center"
  )
```

Describe patterns or conclusions based on this graph.

<br />

## Problem 3

##### Load and tidy the dataset

```{r}
data("ny_noaa")
ny_noaa =
  mutate_at(ny_noaa, vars(date), as.character) %>%
  separate(date, into = c("year", "month", "day"))
```






